k-Space Domain Parallel Transmit Pulse Design

### 2020-09-07 Manuscript/Method.tex
wag methods and figure edits

\section*{Methods}
\subsection*{Design and Simulation Setup}
Simulations were performed to validate and characterize the proposed k-space domain parallel transmit pulse design method,
with comparisons to the spatial-domain method of Ref. \cite{Grissom:2006:MRM}.
RF pulses were designed for a simulated 24-channel loop transmit array (Figure \ref{fig:Coil}) that is under construction for a 7 Tesla scanner optimized
for imaging the human cortex.  
24-channel complex-valued $B_1^+$ maps used in the pulse designs were simulated in a male human head model using 
Ansys High Frequency Structure Simulator (Canonsburg, PA, USA) with 1.5 mm isotropic resolution. 
All pulse designs used a SPINS excitation k-space trajectory \cite{malik2012tailored} (Figures \ref{fig:Target}b\&c) with 5 mm max resolution. 
The SPINS trajectories were first designed analytically and the final gradient and excitation k-space trajectories were designed from them 
using the minimum-time gradient waveform design method \cite{lustig2008fast}, 
for a 15 {\textmu}s dwell time and subject to the scanner's gradient amplitude and slew rate constraints of 200 mT/m and 700 T/m/s, respectively. 
The SPINS trajectory comprised three segments, with radii ranging between 1-0.625, 0.625-0.375, and 0.375-0 /cm, respectively.
The number of polar and azimuthal rotations were 15.5/2.125, 20.66/2.833, 15.5/2.125, respectively, for each segment.
The durations the segments after minimum-time gradient design were 4.174, 4.326, and 1.5 ms, respectively.
With the exception of the simulations across undersampling factors, all pulse designs used the 10 ms trajectory shown in Figures \ref{fig:Target}b\&c.

@@

@@

\par Figure \ref{fig:Target}a shows the target pattern for all pulse designs which comprised an ellipse centered on the ventricles with AP/HF/LR semi-axes of 4.8/3.2/3.2 cm.
The pattern was smoothed by a Fermi filter that was applied in the frequency domain to match the target pattern's effective resolution to the excitation k-space trajectory's 5 mm resolution.
This choice of target pattern was motivated by imaging applications for the 7 Tesla scanner 
in which midbrain signals will be saturated for high-resolution, highly-accelerated imaging of the cortex. 
Pulses were designed to excite the entire ellipse and achieve zero excitation in voxels in the cerebrum but outside the ellipse. 
The smoothed target pattern and the $B_1^+$ maps were downsampled from their original 128$\times$128$\times$96 grids (1.5 mm isotropic resolution) to 
64$\times$64$\times$48 grids (3 mm isotropic resolution),
and RF designs were performed with the 64$\times$64$\times$48 grid size. 
All pulse designs and simulations were performed in MATLAB (Mathworks, Natick, MA, USA).
For spatial domain designs, 
RF pulses were solved using an iterative least-squares conjugate-gradient descent method \cite{Grissom:2006:MRM} with 35 iterations,
which was accelerated by non-uniform fast Fourier transforms \cite{Fessler:2003fk}. 
The proposed k-space domain algorithm was implemented based on code for the non-Cartesian GRAPPA method of Ref. \cite{luo2019grappa},
as a function that takes as input the $B_1^+$ maps, 
a normalized excitation k-space trajectory, and optional off-resonance and design parameters. 
Within that function the Fourier transforms of the $B_1^+$ products are calculated,
and a C-based MEX function is invoked to solve for the non-zero elements of each column of $\bm{W}$. 
Those elements are then inserted into a sparse $\bm{W}$ matrix, which is the single output of the main function. 
Parallelization was implemented across patches within the C-based MEX function using OpenMP. 
The final pulses are calculated by multiplying the sparse $\bm{W}$ matrix into the Fourier transform of the target pattern. 
This code, the $B_1^+$ maps, and a demo script are available at https://github.com/wgrissom/kpTx. 
All pulse designs were performed on a server (Colfax International, Santa Clara, CA, USA) 
with 512 GB RAM and two 24-core 2.1 GHz Intel Xeon CPUs which provide up to 94 threads (Intel Corporation, Santa Clara, CA, USA). 
Designs were performed five times for each case, and the mean computation time was recorded.
For the k-space domain designs, the design time comprised the time to compute the sparse $\bm{W}$ matrix, 
and the time to multiply it into the target pattern to obtain the pulses.
The resulting pulses were Bloch equation-simulated and compared to the target pattern on the finer 128$\times$128$\times$96 grid to capture Gibbs ringing. 
When calculating excitation errors, the magnitude root-mean-square error (RMSE) was calculated in voxels within the cerebrum,
except for an $\approx$5 mm-thick transition band around the edge of the elliptical target region.

@@

>>>>

%!TEX root = kPtx_paper.tex
\section*{Methods}
\subsection*{Design and Simulation Setup}
Simulations were performed to validate and characterize the proposed k-space domain parallel transmit pulse design method,
with comparisons to the spatial-domain method of Ref. \cite{Grissom:2006:MRM}.
RF pulses were designed for a simulated 24-channel loop transmit array (Figure \ref{fig:Coil}) that is being built for a 7 Tesla scanner optimized
for imaging the human cortex.  
24-channel complex-valued $B_1^+$ maps used in the pulse designs were simulated in a male human head model using 
Ansys High Frequency Structure Simulator (Canonsburg, PA, USA) with 1.5 mm isotropic resolution. 
Figure \ref{fig:Target}a shows the target pattern for all pulse designs which comprised an ellipse centered on the ventricles with AP/HF/LR semi-axes of 4.8/3.2/3.2 cm.
The pattern was smoothed by a Fermi filter that was applied in the frequency domain to match the target pattern's effective resolution to the excitation k-space trajectory's 5 mm resolution.
This choice of target pattern was motivated by imaging applications for the 7 Tesla scanner 
in which midbrain signals will be saturated for high-resolution, highly-accelerated imaging of the cortex. 
Pulses were designed to excite the entire ellipse and achieve zero excitation in voxels in the cerebrum but outside the ellipse. 
The smoothed target pattern and the $B_1^+$ maps were downsampled from their original 128$\times$128$\times$96 grids (1.5 mm isotropic resolution) to 
64$\times$64$\times$48 grids (3 mm isotropic resolution),
and RF designs were performed with the 64$\times$64$\times$48 grid size.

@@

\par A SPINS excitation k-space trajectory \cite{malik2012tailored} (Figure \ref{fig:Target}b) with 5 mm max resolution was used for the designs. 
The SPINS trajectory comprised three segments, with radii ranging between 1-0.625, 0.625-0.375, and 0.375-0 /cm, respectively.
The number of polar and azimuthal rotations were 15.5/2.125, 20.66/2.833, 15.5/2.125, respectively, for each segment.
The SPINS trajectories were first designed analytically and the final gradient waveforms and excitation k-space trajectories were designed from them 
using the minimum-time gradient waveform design method \cite{lustig2008fast}, 
for a 15 {\textmu}s dwell time and subject to the scanner's gradient amplitude and slew rate constraints of 200 mT/m and 700 T/m/s, respectively (Figure \ref{fig:Target}c). 
The durations of the three segments after minimum-time gradient design were 4.174, 4.326, and 1.5 ms, respectively.
With the exception of the simulations across undersampling factors, 
all pulse designs used this 10 ms trajectory.

@@

\par All pulse designs and Bloch equation simulations were performed in MATLAB (Mathworks, Natick, MA, USA).
For spatial domain designs, 
RF pulses were solved using an iterative least-squares conjugate-gradient descent method \cite{Grissom:2006:MRM} with 35 iterations,
which was accelerated by non-uniform fast Fourier transforms \cite{Fessler:2003fk}. 
The proposed k-space domain algorithm was implemented based on code for the non-Cartesian GRAPPA method of Ref. \cite{luo2019grappa},
as a function that takes as input the $B_1^+$ maps, 
a normalized excitation k-space trajectory, and optional structures of algorithm and off-resonance parameters. 
Within that function the Fourier transforms of the $B_1^+$ products are calculated,
and a C-based MEX function is invoked to solve for the non-zero elements of each column of $\bm{W}$,
with linear interpolation of the Fourier transforms of the $B_1^+$ map products. 
Those elements are then inserted into a sparse $\bm{W}$ matrix, which is the single output of the main function. 
Parallelization was implemented across patches within the C-based MEX function using OpenMP. 
The final pulses are calculated by multiplying the sparse $\bm{W}$ matrix into the Fourier transform of the target pattern. 
This code, the $B_1^+$ maps, and a demo script are available at https://github.com/wgrissom/kpTx. 
All pulse designs were performed on a server (Colfax International, Santa Clara, CA, USA) 
with 512 GB RAM and two 24-core 2.1 GHz Intel Xeon CPUs which provide up to 94 threads (Intel Corporation, Santa Clara, CA, USA). 
Designs were performed five times for each case, and the mean computation time was recorded.
For the k-space domain designs, the design time comprised the time to compute the sparse $\bm{W}$ matrix, 
and the time to multiply it into the target pattern to obtain the pulses.
The resulting pulses were Bloch equation-simulated and compared to the target pattern on the finer 128$\times$128$\times$96 grid to capture Gibbs ringing. 
When calculating excitation errors, the magnitude root-mean-square error (RMSE) was calculated in voxels within the cerebrum,
except for an $\approx$5 mm-thick transition band around the edge of the elliptical target region.

@@

\subsection*{L-Curves}
To compare the tradeoff between excitation error and integrated RF power for spatial and k-space domain designs,
pulse designs were repeated while varying Tikhonov regularization parameters ($\lambda$ in Equation \ref{eq:Solution}) 
over five orders of magnitude. 
The k-space domain designs were repeated four times to investigate the three main sources of error:
finite patch width, finite inclusion width, and $B_1^+$ map product interpolation. 
Specifically, a design was performed using patch and inclusion widths of four
with interpolated $B_1^+$ map products (`Interpolated $\bm{S}^H\bm{S}$ matrices'),
patch and inclusion widths covering the entire design grid with interpolated $B_1^+$ map products,
patch and inclusion widths of four without $B_1^+$ map product interpolation (`Exact $\bm{S}^H\bm{S}$ matrices'; 
$B_1^+$ maps were phase-modulated to each trajectory location before Fourier transform),
and patch and inclusion widths covering the entire design grid without $B_1^+$ map product interpolation. 
Note that the last case is equivalent to the original k-space domain method of Ref. \cite{Katscher:2003:Magn-Reson-Med:12509830}.

### 2020-09-07 Manuscript/Result.tex
wag methods and figure edits

\section*{Results}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{ComputationTime_2}
	\caption{\textcolor{red}{WAG: Can you also run $2^0$ (most importantly) and $2^5$ patch width here? It would be good to show that the RMSE levels off at some point. Also can you change the RMSE axis to go from 0 to 6? That will help emphasize that even though the error is decreasing with increasing patch width, 
	the difference between patch width = 2 and patch width = something large is quite small. }
	Studies of different parallelization parameters, where patch width is the square root of the number of $\mathbf{W}$ columns solved for simultaneously and inclusion width is the assumed range of influence of each excitation k-space location. \textbf{(a)} Computation time vs. thread numbers. Patch width and inclusion with were kept at 4 cycle/FOV. \textbf{(b\&c)} Computation time (blue axis) and RMSE (red axis) vs. patch widths (b) and inclusion widths (c). Thread number were kept at 16 of for both design. Inclusion width was constant at 4 in (b) and patch widths were constant at 4 (solid line) and 8 (dashed line).}
	\label{fig:ComputationTime}
\end{figure}
Figure \ref{fig:ComputationTime}a shows the mean computation time versus different thread numbers. The k-space design problem held the patch and inclusion width constant at 4. The mean computation time decreases with increasing thread numbers before 12 threads, and then plateaus due to overhead. The thread number was conservatively chosen to be 16 for the following studies to ensure timely solutions. 
Figure \ref{fig:ComputationTime}b shows the mean computation time and the RMSE with different patch widths. The computation time increases and the RMSE decreases as the patch width increases. This was because when patch width increases, there are more excitation trajectory points included in one instance problem, even with the same inclusion width. As a result, each instance problem becomes bigger and more time consuming, which cancels the advantageous fact that there are fewer instance problems in total, especially in cases with parallel computing. Furthermore, increasing the patch width is equivalently increasing the inclusion widths for the center target points within each patch, which makes the solutions more accurate. Here, we chose patch widths 4 and 8 (labeled $2^2$ and $2^3$) to balance time and accuracy for further evaluation.
Figure \ref{fig:ComputationTime}c shows the mean computation time and the RMSE with different inclusion widths. The solid lines and dashed lines were obtained with patch width 4 and 8, respectively. Computation time increases and error decreases with increasing inclusion width. Particularly, increasing inclusion width is equivalent to utilizing more accurate $B_1^+$ information by truncating the $B_1^+$ map Fourier transforms to zero further out from DC. For both patch widths 4 and 8, 4 is the optimal inclusion width, which confirms the result in \ref{fig:ComputationTime}a , previously. Using an inclusion width of 4, the computation time and error are acceptable for both patch width numbers. Following these results,  we chose patch width 4 for the further designs.

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{ErrorMap_2}
	\caption{\textcolor{red}{WAG: I think this is the figure that your example in Github should replicate.}
	Normalized excitation patterns (top row) and error maps (bottom row) for k-space domain design (left column) and spatial domain design (right column), with RMSE and design time indicated.}
	\label{fig:ErrorMap}
\end{figure}
The simulated excitation patterns and error maps are shown in Fig \ref{fig:ErrorMap} for both k-space domain design (left) and spatial domain design (right), where the k-space domain design was solved with 16 parallel computing threads, patch width 4, and inclusion width 4. 
Both designs were done with the 64$\times$64$\times$48 grid size (3mm iso-resolution), and the excitation patterns were evaluated against the target pattern with the 128$\times$128$\times$96 grid size (1.5mm iso-resolution).  grid size. 
The calculated RMSE from the k-space domain design and the spatial domain design were 2.42\% and 2.68\%, respectively, which were practically equal. For both design methods, most of the errors appeared at the edges of the transition band, with other errors lower than 5\% of the target flip angle. This indicates uniform inner volume suppression while maintaining the outer volume intact, which can in turn decrease g-factor in highly accelerated imaging. In practice, error near the transition band can be mitigated by ensuring the region of interest (ROI) of the cortex falls completely in the stop-band and assuming zero signal within the pass-band during undersampled reconstruction. More significantly, the parallelized k-space domain design method drastically reduced the mean computation time from 100.9 seconds to 2.9 seconds compared to the spatial domain design, which was a 97\% decrease.

>>>>

%!TEX root = kPtx_paper.tex
\section*{Results}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{ErrorMap_2}
	\caption{\textcolor{red}{WAG: I think this is the figure that your example in Github should replicate.}
	Normalized excitation patterns (top row) and error maps (bottom row) in central axial, sagittal and coronal slices 
	for k-space domain (left) and spatial domain designs (right).}
	\label{fig:ErrorMap}
\end{figure}
Figure \ref{fig:ErrorMap} shows simulated excitation patterns and error maps for spatial and k-space domain designs. 
The k-space design was performed with 16 parallel threads and patch and inclusion widths of 4. 
Both designs were performed with a 64$\times$64$\times$48 grid size (3 mm isotropic resolution), 
and the excitation patterns were evaluated against the target pattern on a 128$\times$128$\times$96 grid size (1.5 mm iso-resolution). 
The calculated root-mean-squared errors (RMSEs) were 2.42\% (spatial domain) and 2.68\% (k-space domain), 
respectively. 
For both design methods, most of the errors appeared at the edges of the transition band, and errors elsewhere were lower than 5\% of the target flip angle. 
This indicates uniform inner volume excitation while maintaining the outer volume intact. 
The parallelized k-space domain design required 2.9 seconds computation versus 100.9 seconds for the spatial domain method, a 97\% decrease.

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{ComputationTime_2}
	\caption{(a) k-Space design time versus number of threads, for patch and inclusion widths of 4. 
	(b) Studies of different parallelization parameters, where patch width is the square root of the number of $\mathbf{W}$ columns solved for simultaneously and inclusion width is the assumed range of influence of each excitation k-space location. \textbf{(a)} Computation time vs. thread numbers. Patch width and inclusion with were kept at 4 cycle/FOV. \textbf{(b\&c)} Computation time (blue axis) and RMSE (red axis) vs. patch widths (b) and inclusion widths (c). Thread number were kept at 16 of for both design. Inclusion width was constant at 4 in (b) and patch widths were constant at 4 (solid line) and 8 (dashed line).}
	\label{fig:ComputationTime}
\end{figure}
Figure \ref{fig:ComputationTime}a shows the mean computation time versus different thread numbers. The k-space design problem held the patch and inclusion width constant at 4. The mean computation time decreases with increasing thread numbers before 12 threads, and then plateaus due to overhead. The thread number was conservatively chosen to be 16 for the following studies to ensure timely solutions. 
Figure \ref{fig:ComputationTime}b shows the mean computation time and the RMSE with different patch widths. The computation time increases and the RMSE decreases as the patch width increases. This was because when patch width increases, there are more excitation trajectory points included in one instance problem, even with the same inclusion width. As a result, each instance problem becomes bigger and more time consuming, which cancels the advantageous fact that there are fewer instance problems in total, especially in cases with parallel computing. Furthermore, increasing the patch width is equivalently increasing the inclusion widths for the center target points within each patch, which makes the solutions more accurate. Here, we chose patch widths 4 and 8 (labeled $2^2$ and $2^3$) to balance time and accuracy for further evaluation.
Figure \ref{fig:ComputationTime}c shows the mean computation time and the RMSE with different inclusion widths. The solid lines and dashed lines were obtained with patch width 4 and 8, respectively. Computation time increases and error decreases with increasing inclusion width. Particularly, increasing inclusion width is equivalent to utilizing more accurate $B_1^+$ information by truncating the $B_1^+$ map Fourier transforms to zero further out from DC. For both patch widths 4 and 8, 4 is the optimal inclusion width, which confirms the result in \ref{fig:ComputationTime}a , previously. Using an inclusion width of 4, the computation time and error are acceptable for both patch width numbers. Following these results,  we chose patch width 4 for the further designs.

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{L_curve}
	\caption{L-curves for spatial domain and k-space domain pulse designs, 
	repeated across five orders of magnitude of the methods' Tikhonov regularization parameters.
	The k-space domain designs were repeated using patch and inclusion widths of four versus solving for the entire domain at once (`Path / Inclusion Width = 4' versus `Full Patch'),
	and using $B_1^+$ map product interpolation versus phase modulation to each trajectory location (`Interpolated $\bm{S}^H\bm{S}$ matrices' versus
	`Exact $\bm{S}^H\bm{S}$ matrices').}
	\label{fig:LCurves}
\end{figure}
